The Weakly Nonlinear Magnetorotational Instability in a Global, Cylindrical Taylor-Couette Flow

### 2017-03-25 paper/thingap_ref_response/response_to_referee.tex

\item With regard to saturation through modification of the base flow, in the TC problem studied in these papers, the rotation law is enforced through viscous coupling of the interior fluid to the bounding cylinders. This introduces a timescale, the viscous diffusion time, with which the momentum transport driven by perturbations must compete. In an accretion disk the rotation law is enforced at every point directly by gravity. The authors should comment on how important they think this difference is to saturating the mode.

\textbf{In Paper II, the rotation law is enforced through viscous coupling to the boundaries. In Paper I, however, the rotation law is imposed directly at every point by the shear. We have noted this point in the text. See also the newly extended discussion of the local flow with shear periodic boundary conditions.}

\item The influence of boundary conditions and boundary layers is strong. To call out some examples: In Paper I, the amplitude $\alpha$ converges to the constant value $(b/c)^{1/2}$, independent of z. Is this really relevant to a lab experiment with vertical endcaps, or to a stratified disk? Have the authors experimented with a finite cylinder, even a very tall one? Likewise, the radial boundaries seem to play a role that a softer physical boundary might not, for example the boundary layers in Figure 2 of Paper II (in this case it would be useful to show blowups of the solution, as in Figure 4, and present evidence for numerical convergence). I would like to see more discussion of the role of these boundary conditions versus physical boundary conditions. Again, it is OK to model a lab experiment – but the connection to a natural system should be kept in mind.

\textbf{The amplitude $\alpha$ is bounded by the value +/- $(b/c)^{1/2}$, although the amplitude does not necessarily converge to a constant value, as shown in Figures 6 and 7. We note, however, that although the value of the quantity $(b/c)^{1/2}$ is heavily influenced by the boundary layers in the walled TC flow, this is specific to the wall-like boundary conditions and not inherent to the theory. We have added a paragraph at the end of section 4 to elaborate on this point, discuss what the $(b/c)^{1/2}$ saturation amplitude means physically, and describe how the derived amplitude applies to the shearing box MRI. We also confirmed numerical convergence in the widegap geometry and added a sentence noting convergence in the text.}

@@

\item Both papers are light on physical interpretation. This is especially an issue with Paper I -- see below. I would particularly like to see:

\begin{itemize}						 							
\item A more extended discussion of how angular momentum transport works. After all, this is the primary raison d’etre for most studies of the MRI.

\textbf{We have added a plot showing the morphology of the stress in the x-z domain (Fig 5), a comparison of angular momentum transport by Maxwell stresses, Reynolds stresses, and total stresses in both the weakly nonlinear theory and direct numerical simulations (Fig 6), and more discussion both in the "Discussion" section and throughout the text.}
%\textbf{We have added a figure to the thin-gap paper showing the morphology of the average angular momentum transport $\dot{J}$. }

@@

\item A small point – the parameter called $\beta$ is not the standard definition; why not introduce the Cowling number?

\textbf{We have made this change.}

\item Finally, would it be feasible to a compute a nonlinear solution with Dedalus (or some other platform) and see how accurate the perturbation theory really is, for different values of $\epsilon$? Although the authors compare their results with Ebrahimi 2009, it would be desirable to have their own simulations for comparison.

%\textbf{We have carried out a direct numerical situation of the thin-gap TC flow with Dedalus and have added a new section (``Simulations") discussing the results.}

@@

\end{enumerate}

Remarks on Paper I:

This paper follows URM07 very closely – how closely I did not appreciate until I looked at that paper. Unless the authors can substantially enhance their discussion of the physics, bring out details of how their approach differs from URM07 that escaped me, and/or more fully discuss the relationship between their work and other work on saturation such as that of Vasil 2015, Ebrahimi 2009, or the Hatch/Makwana/Terry approach mentioned above, it might be better to write one long paper that outlines the URM07 approach and applies it to the helical MRI. A secondary reason to suggest this is the strong overlap of the Introductions to Papers I and II.

\textbf{A note on another change in Paper I: we realized that the parameterization of destabilization into the weakly nonlinear regime was incorrect in the previously submitted version of this paper, and in URM07: the background field must be raised, rather than lowered, by a factor of $\epsilon^2$ in order to destabilize the system. This is a facet of the nonideal MRI that is treated incorrectly in previous works, and we add text and a figure (Figure 4) to explain this point. We have also amended the appendices of both papers to reflect the correct tuning.  }

\textbf{We thank the referee again for useful comments and probing questions. We hope that the referee will agree that with the extensive additions, including a new section discussing the applicability of the weakly nonlinear theory to the shearing box, the addition of a direct numerical simulation, and three new plots, that these papers are suitable for publication in The Astrophysical Journal. }

### 2017-03-26 paper/thingap_ref_response/response_to_referee.tex

\end{enumerate}

Remarks on Paper I:

This paper follows URM07 very closely – how closely I did not appreciate until I looked at that paper. Unless the authors can substantially enhance their discussion of the physics, bring out details of how their approach differs from URM07 that escaped me, and/or more fully discuss the relationship between their work and other work on saturation such as that of Vasil 2015, Ebrahimi 2009, or the Hatch/Makwana/Terry approach mentioned above, it might be better to write one long paper that outlines the URM07 approach and applies it to the helical MRI. A secondary reason to suggest this is the strong overlap of the Introductions to Papers I and II.

\textbf{A note on another change in both papers: we realized that the parameterization of destabilization into the weakly nonlinear regime was incorrect in the previously submitted version of this paper, and in URM07: the background field must be raised, rather than lowered, by a factor of $\epsilon^2$ in order to destabilize the system. This is a facet of the nonideal MRI that is treated incorrectly in previous works, and we add text and a figure (Figure 4) to explain this point in Paper I. We have also amended the appendices of both papers to reflect the correct tuning. Thanks to a fortuitous sign error the plots were correct for everything but the helical MRI saturated perturbation structure (Paper II, Figure 8), which has been corrected. (The correct parameterization changes the overall saturation amplitude derived from the weakly nonlinear theory, but not the character of the perturbation structure.) }

### 2017-03-26 paper/widegap_ref_response/widegap_mri_paper_ref_response_apj.tex
merged.
fixed acknowledgments

\begin{figure*}
\centering
\includegraphics[width=\textwidth]{widegap_vel_Bfield_streamfuncs_512_final_wider-eps-converted-to.pdf}
\caption{Perturbation structure for the velocity and magnetic field of the fiducial standard MRI case, including first and second order perturbations. Leftmost panel is a radially zoomed-in section of the velocity perturbation structure, to better show the boundary layer-driven structure at the inner cylinder. Colors are azimuthal velocity and magnetic field perturbations, and streamfunctions show the perturbation structure in the $r, z$ plane. The width of the streamfunctions is proportional to the speed and magnetic field strength in the $r, z$ plane for the velocity and magnetic field, respectively. Vertical domain covers one critical wavelength $\lambda_c = 2\pi / k_c$. We use the constant saturation amplitude $\alpha_{s} = 3.9 \times 10^{-5}$ derived for this case, and a small parameter $\epsilon = 0.5$.}\label{fig:smri_streamfuncs}
\end{figure*}

@@

\begin{figure*}
\centering
\includegraphics[width=\textwidth]{hmri_velocity_Bfield_streamfuncs-eps-converted-to.pdf}
\caption{As in Figure \ref{fig:smri_streamfuncs} but for the fiducial helical MRI case, including first and second order perturbations.}
\end{figure*}

@@

\section{Acknowledgments}
S.E.C. was supported by a National Science Foundation Graduate Research Fellowship under grant No. DGE-16-44869. J.S.O. acknowledges support from NASA grant 15-LWS15\_2-0063. The authors thank Mordecai Mac Low, Jeremy Goodman, John Krommes, Geoff Vasil, and Ellen Zweibel for useful discussion.
\software{Dedalus: http://dedalus-project.org/}

>>>>

\begin{figure*}
\centering
\includegraphics[width=\textwidth]{widegap_vel_Bfield_streamfuncs_512_final_wider}
\caption{Perturbation structure for the velocity and magnetic field of the fiducial standard MRI case, including first and second order perturbations. Leftmost panel is a radially zoomed-in section of the velocity perturbation structure, to better show the boundary layer-driven structure at the inner cylinder. Colors are azimuthal velocity and magnetic field perturbations, and streamfunctions show the perturbation structure in the $r, z$ plane. The width of the streamfunctions is proportional to the speed and magnetic field strength in the $r, z$ plane for the velocity and magnetic field, respectively. Vertical domain covers one critical wavelength $\lambda_c = 2\pi / k_c$. We use the constant saturation amplitude $\alpha_{s} = 3.9 \times 10^{-5}$ derived for this case, and a small parameter $\epsilon = 0.5$.}\label{fig:smri_streamfuncs}
\end{figure*}

@@

\begin{figure*}
\centering
\includegraphics[width=\textwidth]{hmri_velocity_Bfield_streamfuncs_correct_satamp}
\caption{As in Figure \ref{fig:smri_streamfuncs} but for the fiducial helical MRI case, including first and second order perturbations.}
\end{figure*}

@@

\section{Acknowledgments}
S.E.C. was supported by a National Science Foundation Graduate Research Fellowship under grant No. DGE-16-44869. J.S.O. acknowledges support from NASA grant \edit1{NNX16AC92G}. \edit1{We thank the anonymous referee for thoughtful comments. We also thank} Mordecai Mac Low, Jeremy Goodman, John Krommes, Geoff Vasil, and Ellen Zweibel for useful discussion.
\software{Dedalus: http://dedalus-project.org/}
